mathml: Translate R expressions to MathML and LaTeX/MathJax

### 2022-12-21 Draft3.Rmd
fix problems with parentheses

term <- quote(NaN + NA + TRUE + FALSE + Inf + -Inf)
mathout(term)
```

@@

Note that for the standard operators with known precedence, \pkg{mathml} is able
to detect if parentheses are needed; for example, a parenthesis is automatically
put around `g + h` in the `nodot`-example. However, we also note an unecessary
parenthesis around `l + m` above. This parenthesis is a consequence
of `quote(a^(b + c))` actually producing a nested R\ call of the
form `'^'(a, (b + c))` instead of `'^'(a, b + c)`:

@@

For self-written functions, matters are a bit more complicated because the
underlying R package `rolog` [@rolog] cannot easily access the function name,
hence a default is chosen.

```{r, results='asis'}
f <- function(x)
{
  if(x == 0L) return(0L);
  if(x <= 0L) return(-1L);
  if(x > 0L) return(1L)
}

Custom names can be given via attributes.

```{r, results='asis'}
attributes(f)$name = quote(sgn)
mathout(f)
```

@@

One of R's great features is the possibility to refer to function arguments by
by their names, not only by their position in the list of arguments.
Unfortunately, the prolog-powered handler for `integrate/3` is rather rigid; it
accepts only three arguments in that particular order and without names, that
is, `integrate(lower=0L, upper=2L*pi, sin)`, would not print the desired result.

To "canonicalize" function calls with named arguments and arguments in unusual
order, \pkg{mathml} provides an auxilliary function `canonical(f, drop)` that
reorders the argument list of calls to known functions and, if `drop=TRUE`,
drops the function names.

@@

```{r, echo=TRUE, results='asis'}
# A custom function
g <- function(u)
{
  sin(u)
}

# Mixture of (partially) named and positional arguments in unusual order
term = call("integrate", lower=-Inf, upper=Inf, g)
attributes(term)$name = quote(g)
mathout(canonical(term))

# Even more complex
term <- call("*", 2L, call("$", call("integrate", low=-Inf, up=Inf, g), "value"))
attributes(term)$name = quote(g)
mathout(canonical(term))

>>>>

term <- quote(NaN + NA + TRUE + FALSE + Inf + (-Inf))
mathout(term)
```

@@

For the standard operators with known precedence, \pkg{mathml} is generally able
to detect if parentheses are needed; for example, parentheses are automatically
put around `g + h` in the `nodot`-example. However, we note unecessary
parentheses around `l + m` above. This is parenthesis is a consequence
of `quote(a^(b + c))` actually producing a nested R\ call of the
form `'^'(a, (b + c))` instead of `'^'(a, b + c)`:

@@

For self-written functions, matters are a bit more complicated because the
underlying R package `rolog` [@rolog] cannot easily access the function name,
hence a default is chosen. The function name can be given as an attribute.

```{r, results='asis'}
f <- function(x)
{
  if(x == 0L) return(0L);
  if(x < 0L) return(-1L);
  if(x > 0L) return(1L)
}

attributes(f)$name = quote(sgn)
mathout(f)
```

@@

One of R's great features is the possibility to refer to function arguments by
their names, not only by their position in a list of arguments. Unfortunately, 
the Prolog-powered handler for `integrate/3` is rather rigid; it accepts only
three arguments in that particular order and without names, that
is, `integrate(lower=0L, upper=2L*pi, sin)`, would not print the desired result.

To "canonicalize" function calls with named arguments and arguments in unusual
order, \pkg{mathml} provides an auxilliary function `canonical(f, drop)` that
reorders the argument list of calls to known functions and, if `drop=TRUE`, also
removes the names of the arguments.

@@

We can then easily feed mixtures of partially named and positional arguments in
unusual order into the renderer.

```{r, echo=TRUE, results='asis'}
g <- function(u)
{
  sin(u)
}

# 
term <- call("integrate", lower=-Inf, upper=Inf, g)
attributes(term)$name = quote(g)
mathout(canonical(term))

# 2
term <- call("*", 2L, call("$", call("integrate", low=-Inf, up=Inf, g), "value"))
attributes(term)$name = quote(g)
mathout(canonical(term))
